---
title: "Muusoctopus Tank Activity"
author: "Cheyne Springbett"
date: "`r Sys.Date()`"
output: html_document
---


```{r}
library(remotes)
#install_github("KirtOnthank/CircularTimeHistogram")
library(CircularTimeHistogram)
library(googlesheets4)
library(maptools)
library(png)
library(lubridate)
```

This block of code reads in data from a sheet on a google sheet titled "Tank Burrows" using the read_sheet function from the googlesheets4 library.
```{r}
activity=read_sheet("1ckVjS4Jw9o_eg_l_TAwDFNp_evuu36sFHmXyUqT6YqA", sheet="Tank Burrows")
```


This block of code processes the data by replacing parts of the "File Name" column with new values, in order to create a new column "time". Then, it creates another column "realtime", by parsing the "time" column as a date-time object using the strptime function.
```{r}
activity$time=gsub("\\d{5}__TE_D(\\d{4})(\\d{2})(\\d{2})_T(\\d{2})(\\d{2})(\\d{2}).*","\\1-\\2-\\3 \\4:\\5:\\6", activity$`File Name`)

activity$realtime=strptime(activity$time,"%Y-%m-%d %H:%M:%S")
```

## Filtering out the distrubed/idle triggers
This block of code filters out any rows where the "Filter" column is not "Normal".
```{r}
activity=activity[activity$Filter=="Normal",]
```



This block of code creates a table showing the frequency of each value in the "Filter" column.
```{r}
table(activity$Filter)
```

# Plots
The next five blocks of code use the DayHist function from the CircularTimeHistogram library to create plots of octopus activity data, filtered by behavior and tank. The date parameter is set to "07/31/2022" and the hist.zoom parameter is set to different values for each plot.
```{r}
png("entering.png")
DayHist(activity$realtime[activity$Behavior=="Entering Burrow"],date="07/31/2022",hist.zoom=4.5)
dev.off()
```

```{r}
png("exiting.png")
DayHist(activity$realtime[activity$Behavior=="Exiting Burrow"],date="07/31/2022",hist.zoom=4)
dev.off()
```


```{r}
png("total_acitivty.png")
DayHist(activity$realtime,date="07/31/2022",hist.zoom=12)
dev.off()
```


```{r}
png("burrowers_total_acivity.png")
DayHist(activity$realtime[activity$Octopus_Tank!="Amanda_CS2"&
                            activity$Octopus_Tank!="Joe_CS1"&
                            activity$Octopus_Tank!="Naddlie_CS1"&
                            activity$Octopus_Tank!="Rian_CS5"],
        date="07/31/2022",
        hist.zoom=9)
dev.off()
```


```{r}
png("nonburrowers_total_activity.png")
DayHist(activity$realtime[activity$Octopus_Tank=="Amanda_CS2"|activity$Octopus_Tank=="Joe_CS1"|activity$Octopus_Tank=="Naddlie_CS1"|activity$Octopus_Tank=="Rian_CS5"],
        date="07/31/2022",
        hist.zoom=13)
dev.off()
```


The final block of code creates a table showing the frequency of each combination of behavior and tank.
```{r}
table(activity$Behavior,activity$Octopus_Tank)
```


# Statistical Tests

Running a chi-square test on total activity
```{r}
activity.total=c(sum(hour(activity$realtime)>6&hour(activity$realtime)<21,na.rm=T),
  sum(hour(activity$realtime)<6|hour(activity$realtime)>21,na.rm=T))

activity.total.probs=c(15/24,9/24)



activity.total.chi=chisq.test(activity.total,p=activity.total.probs)

activity.total.chi
```



Running a chi-squared test on activity of only burrowing octopuses
```{r}
activity.burrowers=c(
  sum(hour(activity$realtime[activity$Octopus_Tank!="Amanda_CS2"&
                            activity$Octopus_Tank!="Joe_CS1"&
                            activity$Octopus_Tank!="Naddlie_CS1"&
                            activity$Octopus_Tank!="Rian_CS5"])>6&
        hour(activity$realtime[activity$Octopus_Tank!="Amanda_CS2"&
                            activity$Octopus_Tank!="Joe_CS1"&
                            activity$Octopus_Tank!="Naddlie_CS1"&
                            activity$Octopus_Tank!="Rian_CS5"])<21,na.rm=T),
  sum(hour(activity$realtime[activity$Octopus_Tank!="Amanda_CS2"&
                            activity$Octopus_Tank!="Joe_CS1"&
                            activity$Octopus_Tank!="Naddlie_CS1"&
                            activity$Octopus_Tank!="Rian_CS5"])<6|
        hour(activity$realtime[activity$Octopus_Tank!="Amanda_CS2"&
                            activity$Octopus_Tank!="Joe_CS1"&
                            activity$Octopus_Tank!="Naddlie_CS1"&
                            activity$Octopus_Tank!="Rian_CS5"])>21,na.rm=T)
  )

activity.total.probs=c(15/24,9/24)



activity.burrowers.chi=chisq.test(activity.burrowers,p=activity.total.probs)

activity.burrowers.chi
```
Chi-squared test on entering burrow events
```{r}
activity.entering=activity$realtime[activity$Behavior=="Entering Burrow"]

activity.entering=c(sum(hour(activity.entering)>6&hour(activity.entering)<21,na.rm=T),
  sum(hour(activity.entering)<6|hour(activity.entering)>21,na.rm=T))

activity.entering.probs=c(15/24,9/24)



activity.entering.chi=chisq.test(activity.entering,p=activity.entering.probs)

activity.entering.chi
```
Chi-squared for exiting events
```{r}
activity.exiting=activity$realtime[activity$Behavior=="Exiting Burrow"]

activity.exiting=c(sum(hour(activity.exiting)>6&hour(activity.exiting)<21,na.rm=T),
  sum(hour(activity.exiting)<6|hour(activity.exiting)>21,na.rm=T))

activity.exiting.probs=c(15/24,9/24)



activity.exiting.chi=chisq.test(activity.exiting,p=activity.exiting.probs)

activity.exiting.chi
```

## Chi Squared of night/day activity distribution of burrowers versus non-burrowers

```{r}
activity.nonburrowers=c(
  sum(hour(activity$realtime[activity$Octopus_Tank=="Amanda_CS2"|
                            activity$Octopus_Tank=="Joe_CS1"|
                            activity$Octopus_Tank=="Naddlie_CS1"|
                            activity$Octopus_Tank=="Rian_CS5"])>6&
        hour(activity$realtime[activity$Octopus_Tank=="Amanda_CS2"|
                            activity$Octopus_Tank=="Joe_CS1"|
                            activity$Octopus_Tank=="Naddlie_CS1"|
                            activity$Octopus_Tank=="Rian_CS5"])<21,na.rm=T),
  sum(hour(activity$realtime[activity$Octopus_Tank=="Amanda_CS2"|
                            activity$Octopus_Tank=="Joe_CS1"|
                            activity$Octopus_Tank=="Naddlie_CS1"|
                            activity$Octopus_Tank=="Rian_CS5"])<6|
        hour(activity$realtime[activity$Octopus_Tank=="Amanda_CS2"|
                            activity$Octopus_Tank=="Joe_CS1"|
                            activity$Octopus_Tank=="Naddlie_CS1"|
                            activity$Octopus_Tank=="Rian_CS5"])>21,na.rm=T)
  )


activity.mat=rbind(activity.burrowers,activity.nonburrowers)
colnames(activity.mat)=c("day","night")
row.names(activity.mat)=c("burrowers","non.burrowers")
activity.mat

prop.test(activity.mat)

```

```{r}
barplot(t(activity.mat))
```

